\chapter{Master Thesis project}
\label{sec:projectOverview}
In the last chapter of the Thesis Preparation report, the project plan will be introduced as well as a broad overview of different 
work packages. Further, the project timeline can be seen as a Gantt chart.
Probably, there are some parts which will not work out as expected and 
adjustments are needed throughout the Thesis, the project plan can be seen as a rough guideline.

\section{Problem conclusion:}
Conclusion from red boxes.

During the Master Thesis, the reconstruction problem with unknown angles is considered. 
Moreover, the observed samples are considered to be noisy. 
The resulting proposed algorithm should work in the 2D and 3D scenario (classical tomography and cryo-Em).

The main idea is to exploit the fact, that the underlying manifold is known (circle in 2D and sphere in 3D). 
From our noisy observations, a manifold can be computed and compare it with the original manifold.
The comparison between the manifolds enables the possibility of a loss function and learning in general.

It is expected, that the folded spectrum \cite{foldedSpectrumMethod} introduced in section~\ref{sec:FoldedSpectrumMethod}
can be used to estimate the eigenvalues of the Graph Laplacian.
Further, as already mentioned in section~\ref{sec:wasserstein-metric}, the wasserstein metric is a good choice
as a loss function when it comes to dealing with data from a manifold distribution \cite{wassersteinGAN}, as in our case. 

The problem can be seen as Graph Denoising as observations are noisy and therefore, the proposed algorithm 
will denoise the graph based on the manifold assumption. 



\paragraph{Evaluation:}
During evaluation, 2D and 3D scenario will be considered. A first evaluation will be done on artificial constructed
toy-dataset. If time allows, real dataset from classical tomography and/or cryo-EM\footnote{https://www.ebi.ac.uk/emdb/} can be evaluated as well.
During evaluation, two baselines are considered which already solved the problem. The first one is a multi-frequency diffusion 
map approach\cite{multiDiffusionMaps, cryoEmMutliDM}, which aims to denoise cryo-EM images. 
Secondly, \cite{LaplaceRandomProjections} a Graph Laplacian approach solving classical tomography with random projection angles will be compare against.
The evaluation process is a first broad idea. Any adjustments in baseline papers or dataset are possible during 
the Master Thesis. The baseline papers are further addressed in the related work chapter~\ref{sec:relatedWork}
and detailed work packages are defined in chapter~\ref{sec:projectPlan}


\section{Work packages}

\paragraph{Implement algorithm for 2D case:}
The first step will be, to familiarize with the problem and implement
the algorithm for 2D. 

\paragraph{Evaluate 2D case on toy dataset and implement baselines:}
As a second step, the implemented 2D algorithm will be tested on a toy dataset,
where noise is added to the images by hand. As the aim is to work with highly noisy images,
the noise level can be selected and increased when working with toy datasets. 
The evaluation in 2D is crucial and needs to in a satisfying matter. 
It does not make sense to continue with 3D implement, when the simply 2D case is not handled well enough.
Therefore, if the evaluation results are not satisfying, the algorithm needs to be iteratively adjusted, 
such that the evaluation will be in a good enough quality.


\paragraph{Implement algorithm for 3D case:}
After successfully evaluating the algorithm in 2D, the aim is to extend the algorithm to work in 3D as well.

\paragraph{Evaluate 3D case on toy dataset and adjust baselines:}
Again, the implementation will be evaluated on a toy dataset, where noise can be adjusted by hand.


\paragraph{Nice to have: Evaluate of real dataset}
If time allows and the 2D and 3D implementation are evaluated successfully on toy datasets, 
real data can be used for further evaluation. This step will only be done, if time allows.


\paragraph{Evaluate related work:}
As cryo-EM reconstruction is a hot research topic, related work can not only
be considered during the start of the Thesis and needs to be evaluated throughout the Thesis.

\paragraph{Writing Thesis:}
Document implementation and evaluation result.

\begin{landscape}

    \section{Gantt chart}

    \newganttchartelement{bluebar}{
    bluebar/.style={
        inner sep=0pt,
        draw=purple!44!black,
        very thick,
        top color=white,
        bottom color=blue!80
    },
    bluebar label font=\slshape,
    bluebar left shift=.1,
    bluebar right shift=-.1
}

\newganttchartelement{greenbar}{
    greenbar/.style={
        inner sep=0pt,
        draw=green!50!black,
        very thick,
        top color=white,
        bottom color=green!80
    },
    greenbar label font=\slshape,
    greenbar left shift=.1,
    greenbar right shift=-.1
}

\ganttset{%
calendar week text={W\currentweek}%
}

    
    \begin{adjustwidth}{-70pt}{}
        \begin{ganttchart}[expand chart=1.8\textwidth,
            hgrid style/.style={black, dashed},
            vgrid={*{6}{draw=none},dotted},
            x unit=3pt,
            y unit chart=20pt,
            time slot format=isodate,
            group label font=\bfseries \Large,
            ]{2021-12-01}{2022-05-29}
            \gantttitlecalendar{year, month=name, week}{1}\\
        
            \ganttgroup[ group/.append ]{2D classical tomography}{2021-12-01}{2022-02-13}\\ 
                \ganttbluebar[name=Implementation 2D]{Implementation in 2D}{2021-12-01}{2022-01-09}\\ 
                \ganttbluebar[name=Implement Baselines]{Implement Baselines for 2D Evaluation}{2022-01-10}{2022-01-23}\\ 
                \ganttbluebar[name=Evaluation 2D]{Evaluation 2D on toy dataset}{2022-01-24}{2022-02-13}\\ 
                \ganttbluebar[name=Related Work]{Related Work}{2022-01-01}{2022-01-10}\\ 
                \ganttgreenbar[name=Honeymoon]{Honeymoon}{2021-12-20}{2022-01-10}
                
            \ganttnewline[thick, black]

            \ganttgroup[ group/.append]{3D cyro-EM}{2022-02-13}{2022-04-10}\\ 
                \ganttbluebar[name=Implementation 3D]{Implementation in 3D}{2022-02-14}{2022-03-06}\\ 
                \ganttbluebar[name=Extend Baselines 3D]{Extend Baselines for 3D Evaluation}{2022-03-07}{2022-03-20}\\ 
                \ganttbluebar[name=Evaluation 3D]{Evaluation 3D on toy dataset}{2022-03-21}{2022-04-10}
            
                \ganttnewline[thick, black]
        
            \ganttgroup[ group/.append ]{Finalization}{2022-04-11}{2022-05-29}\\ 
                \ganttbluebar[name=Final Evaluation Result]{Final Evaluation Results}{2022-04-11}{2022-04-24}\\ 
                \ganttbluebar[name=Thesis]{Writing Thesis}{2022-04-25}{2022-05-29}\\
                \ganttgreenbar[name=Fatherhood]{Birth of first child}{2022-04-20}{2022-04-27}
        
            %Implementing links
            \ganttlink[link bulge=2, link mid=0.5]{Implementation 2D}{Implement Baselines}
            \ganttlink[link bulge=2, link mid=0.5]{Implement Baselines}{Evaluation 2D}
            
            \ganttlink[link bulge=2, link mid=0.5]{Evaluation 2D}{Implementation 3D}
        
            \ganttlink[link bulge=2, link mid=0.5]{Implementation 3D}{Extend Baselines 3D}
            \ganttlink[link bulge=2, link mid=0.5]{Extend Baselines 3D}{Evaluation 3D}
            
            %\ganttlink[link bulge=0.5, link mid=0.01]{Implementation 2D}{Clean-Up}
            %\ganttlink[link bulge=2, link mid=0.3]{Implementation 3D}{Clean-Up}
        
        
            \ganttlink[link bulge=2, link mid=0.5]{Evaluation 2D}{Final Evaluation Result}
            \ganttlink[link bulge=2, link mid=0.5]{Evaluation 3D}{Final Evaluation Result}
            \ganttlink[link bulge=2, link mid=0.5]{Final Evaluation Result}{Thesis}
        
        \end{ganttchart}
        \end{adjustwidth}

\end{landscape}